---
title: "HW4 MD CityCrime"
#Madeline Seck 
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse) 
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}


md_cities_rates <- md_cities %>%
  mutate(
    violent_rate_2019  = (violent2019 / pop2019) * 100000,
    property_rate_2019 = (property2019 / pop2019) * 100000,
    violent_rate_2020  = (violent2020 / pop2020) * 100000,
    property_rate_2020 = (property2020 / pop2020) * 100000
  )

head(md_cities_rates)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

#In both 2019 and 2020, Baltimore has by far the highest violent crime rate that any city in Maryland. Smaller counties such as Salisbury and Hagerstown have a higher crime rate, but a small population. Barnesville has no crime.

#A news story would be Boltimore has the has the top violent crimes while smaller town have higher property crime.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates %>%
  mutate(
    violent_rate_change  = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) * 100,
    property_rate_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100
  )

head(md_cities_change)


```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?
#the percent change is smaller in towns with very few incidents, makeing them outliers. 
#larger cities like Baltimore have a the highest voilent crimw rate, but the yealy percent change is not a lot 

#the story is that Baltimore has the highest violent crime rate in Maryalnd as a whole, smaller places like Salisbury, Hagerstown only have a handfull of crimes 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

#Baltimore continues to have the highest violent crime rate in Maryland  according to state crime data from 2019 and 2020.

#The data show that violent crime in Baltimore is higher than other cities in Maryland, smaller cities such as Salisbury and Hagerstown high propertyr crime rates.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.



#HW 5 
#Madeline Seck

#1 Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile?


```{r}
md_cities_percentiles <- md_cities_rates %>%
  mutate(
    violent_2020_percentile = ntile(violent_rate_2020, 100)  # creates percentiles
  )
above_90 <- md_cities_percentiles %>%
  filter(violent_2020_percentile > 90)

```

#2Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works.
```{r}
md_cities_rates %>%
  summarize(
    violent2020_IQR = IQR(violent_rate_2020, na.rm = TRUE)
  )
```

#3 Do the results change your best story idea from the previous assignment? Why or why not?
The results do not change by story idea, Baltimoree still has the highest violent crime rate in Maryland. The data show more cities such a ocean city, cambride, Princess Anne, etc havinf high propwrty crime bast on their small population size

