---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse) 
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
mean_dance <- mean(spotify$danceability, na.rm = TRUE)
sd_dance <- sd(spotify$danceability, na.rm = TRUE)

```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify %>%
  summarise(
    mean_danceability = mean(danceability, na.rm = TRUE),
    sd_danceability   = sd(danceability, na.rm = TRUE)
  )
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
hist(spotify$danceability,
     main = "Distribution of Danceability Scores",
     xlab = "Danceability",
     col = "lightblue",
     border = "white")
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
mean_dance <- mean(spotify$danceability, na.rm = TRUE)

ggplot(spotify, aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  geom_vline(aes(xintercept = mean_dance), 
             color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distribution of Danceability Scores",
    x = "Danceability",
    y = "Count"
  ) +
  theme_minimal()
```


### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
mean_dance <- mean(spotify$danceability, na.rm = TRUE)
sd_dance   <- sd(spotify$danceability, na.rm = TRUE)

ggplot(spotify, aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  geom_vline(aes(xintercept = mean_dance), 
             color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean_dance - sd_dance), 
             color = "darkgreen", linetype = "dotted", size = 1) +
  geom_vline(aes(xintercept = mean_dance + sd_dance), 
             color = "darkgreen", linetype = "dotted", size = 1) +
  labs(
    title = "Distribution of Danceability Scores",
    x = "Danceability",
    y = "Count"
  ) +
  theme_minimal()
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.

#The distribution of danceability scores is skewed to the right, with a mean of 0.70 and a standard deviation of 0.14. This indicates that most songs are highly danceable, while fewer tracks have lower danceability. A potential story could explore which genres or artists are producing the most danceable music, feature a profile or opinion piece on their work, or cover a concert and include some of this data in the story.

### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
