---
title: "HW7_MandMs"
date: 10/1/2025
---


In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load an M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
url <- "https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv"
head(mandms)
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
set.seed(1234)
sample5 <- sample_n(mandms, 5)
sample10 <- sample_n(mandms, 10)
sample5
sample10

```


### Task 4: Calculate the mean, median and standard deviation of red M&Ms from each of the samples.
```{r}
mean_red_5 <- mean(sample5$red)
median_red_5 <- median(sample5$red)
sd_red_5 <- sd(sample5$red)

mean_red_5
median_red_5
sd_red_5

mean_red_10 <- mean(sample10$red)
median_red_10 <- median(sample10$red)
sd_red_10 <- sd(sample10$red)

mean_red_10
median_red_10
sd_red_10

```


### Task 5: Calculate the mean, median and standard deviation of red M&Ms from the population

```{r}
mean_red_pop <- mean(mandms$red)
median_red_pop <- median(mandms$red)
sd_red_pop <- sd(mandms$red)
mean_red_pop
median_red_pop
sd_red_pop
```

### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Include a headline. Use a consistent binwidth.

```{r}
ggplot(sample5, aes(x = red)) +
  geom_histogram(binwidth = 2, fill = "purple", color = "black") +
  labs(title = "Distribution of Red M&Ms (Sample of 5)",
       x = "Number of Red M&Ms",
       y = "Count")

ggplot(sample10, aes(x = red)) +
  geom_histogram(binwidth = 2, fill = "pink", color = "black") +
  labs(title = "Distribution of Red M&Ms (Sample of 10)",
       x = "Number of Red M&Ms",
       y = "Count")

ggplot(mandms, aes(x = red)) +
  geom_histogram(binwidth = 2, fill = "red", color = "black") +
  labs(title = "Distribution of Red M&Ms (Population)",
       x = "Number of Red M&Ms",
       y = "Count")


```


### Task 7: Answer questions


Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?
#The means and meduans are close to the poluation. The median are slightly higher than the population. 

What does this exercise tell you about the relationship between sample size and estimate accuracy?

#larger samples lead to more reliable and accurate estimates of the population.

How does random sampling introduce variation into statistical estimates?
#Random sampling introduces variation because each sample includes different set of numbers, causing the results to change slightly each time.

If you repeated this process, would the samples always be identical? Why or why not?
#No, because the sameple size is always ramdom 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
