---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
#Madeline Seck 
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

# Teams Rainbows

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**
# 0.5 is 50%, our mean ours is 86%, which is higher than the mean. 
#Diferant groups have different result depending on the number of the M&Ns they have. We each get a ramdom sameple. 
#our group is 86% and others are 90%, our mean is lower 
# A ramdom sample can apply depending on the the demegraphics from different areas 
------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**
# The population is 53,485 on campus and they only surveyed 12,233. They only sample about 23% of the population. It is less than half of the population. It does not provide a good representation of the student body, which was surveyed by only 18%. It was also less representative of minority communities. This survey is missing perspectives. ------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
population <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

head(population)
```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_100 <-population |> 
  sample_n(100)
```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_100 |> summarize(mean = mean(age), median = median(age), sd = sd(age))


```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}

population |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics
**A**
Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

#The means and SD are different. In the sample size, the mean is slightly lower than the medieum.The second dataset is just a little lower on average and slightly more spread out.

**A**:

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
