---
title: "Sampling_Demo"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```



### Task 2: Load 100 rows of data with the ages of people from 12-30
```{r}

set.seed(42)
younguns <- tibble(
  age = sample(12:30, 100, replace = TRUE)
)

head(younguns)

```



### Task 3: Sample 50 or 10 rows of data 
```{r}
sample_50 <-younguns |> 
  sample_n(50)

sample_10 <-younguns |> 
  sample_n(10)
```


### Task 4: Calculate the mean age and standard deviation of the sample
#50 sample
```{r}
sample_50 |> summarize(mean = mean(age), median = median(age), sd = sd(age))
```

#10 sample
```{r}

sample_10 |> summarize(mean = mean(age), median = median(age), sd = sd(age))
```



### Task 5: Calculate the mean age and standard deviation of the population
```{r}
younguns |> summarize(mean = mean(age), median = median(age), sd = sd(age))
```

#Nerd bonus: Skew!

```{r}
younguns |> 
 summarize(
   mean = mean(age), 
   sd = sd(age), 
   median = median(age)) |> 
 mutate(skew = 3*(mean-median)/sd)
```

