---
title: "Final Project "
author: "Madeline Seck "
date: "11-14-25"
---
**"Madeline Seck"**

```{r}
library(tidyverse)
```


```{r}
billboard <- read_csv("billboard.csv")
```

```{r}
view(billboard)
```


#Part 1: Data Explorations
1. Describe a hypothesis about what makes a hit song or come up with some other question/hypothesis.
##Songs with higher overall ratings spend more weeks at number one through the overall time periods. The earliest year is 1958 and lastest year is 2025 and the data includes 1178 songs.

-- What does it contain and what time period does it cover?
##the columns that are mose important my analysis are "Overall Rating" and  "Weeks at Number One."

```{r}
ggplot(billboard, aes(x = `Overall Rating`, y = `Weeks at Number One`)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Relationship Between Overall Rating and Weeks at #1",
    x = "Overall Rating",
    y = "Weeks at #1"
  )

```
```{r}
cor_test <- cor.test(billboard$`Overall Rating`, billboard$`Weeks at Number One`)
print(cor_test)

```
##0.14 shows a weak positive relationship between a song’s overall rating and the number of weeks it spends at number one. The p-value is much less than 0.05. This allows us to reject the null hypothesis that there is no relationship between overall rating and chart success. Though overall rating alone only explains a small portion of a song’s chart longevity. 

--What's missing that you might need or want?
#There is an outlier with a rating of 6. Other factors, like genre, era, promotion or artist popularity, likely play a role.


# For Weeks at Number One
```{r}
mean(billboard$`Weeks at Number One`, na.rm = TRUE)
median(billboard$`Weeks at Number One`, na.rm = TRUE)
min(billboard$`Weeks at Number One`, na.rm = TRUE)
max(billboard$`Weeks at Number One`, na.rm = TRUE)
sd(billboard$`Weeks at Number One`, na.rm = TRUE)

```
# For Overall Rating
```{r}
mean(billboard$`Overall Rating`, na.rm = TRUE)
median(billboard$`Overall Rating`, na.rm = TRUE)
min(billboard$`Overall Rating`, na.rm = TRUE)
max(billboard$`Overall Rating`, na.rm = TRUE)
sd(billboard$`Overall Rating`, na.rm = TRUE)

```

Next, propose a newsworthy hypothesis (or several of them) for your dataset. You'll be looking to find (or eliminate possible story ideas) using the techniques you've learned in this class. You'll start that effort this week by coming up with some theories or questions that you're interested in exploring that can be answered with the dataset you selected.

#hypothesis
#3. What top 3 genres are in the charts longer? I think pop is, hip-hop and country

#alternate
#1. Songs released in more recent decades are in  the charts longer. 
#2. 1. Songs released in older decades are in  the charts longer.


#Part 2 


```{r}
songs <- read.csv("billboard.csv")

```

```{r}
view(billboard)
```


```{r}
library(dplyr)
```

#Top 3 genres with the longest average time at number one.

```{r}
avg_weeks_genre <- df %>%
  group_by(CDR.Genre) %>%
  summarize(mean_weeks = mean(Weeks.at.Number.One, na.rm = TRUE)) %>%
  arrange(desc(mean_weeks))
head(avg_weeks_genre, 3)
```

#anova 

```{r}
df_genres <- data.frame(
  CDR.Genre = c("Electronic/Dance;Latin", "Hip Hop;Reggae", "Pop;Reggae"),
  Weeks = c(14.0, 9.0, 6.5)
)
anova_result <- aov(Weeks ~ CDR.Genre, data = df_genres)
summary(anova_result)
```

#z -score 
```{r}
df$z_score <- (df$`Weeks.at.Number.One` - mean(df$`Weeks.at.Number.One`, na.rm = TRUE)) / 
               sd(df$`Weeks.at.Number.One`, na.rm = TRUE)
head(df[order(-df$z_score), c("Song", "Artist", "Weeks.at.Number.One", "z_score")], 10)

```


#ggplot 
```{r}
ggplot(avg_weeks_genre, aes(x = reorder(CDR.Genre, mean_weeks), y = mean_weeks, fill = mean_weeks)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # horizontal bars for readability
  labs(title = "Average Weeks at #1 by Genre",
       x = "Genre",
       y = "Average Weeks at Number One") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  theme_minimal()
```

#summary 
The dataset contains a total of 1,177 songs. I analyzed which genres remained at number one on the charts the longest. The top genres were Electronic/Dance; Latin, averaging 14 weeks, followed by Hip Hop; Reggae at 9 weeks, and Pop; Reggae at 6.5 weeks. These genres performed well regardless of other factors that might play a role 

#Anova Results 
This analysis confirms that Electronic/Dance; Latin songs tend to remain at number one longer than Pop; Reggae songs. It suggests that the differences between genres are meaningful and not due to random chance.

#Z-score 
To identify the songs that stayed longest on the charts, the top song was the hip-hop hit Old Town Road by Lil Nas X featuring Billy Ray Cyrus, which remained on the charts for 19 weeks and had a z-score of 6.08, marking it as a clear outlier in the dataset.

#What is missing?
A potential test I  could explore which genre(s) are least popular on the charts. Another testable question is whether Hip Hop and Reggae frequently crossover into other genres and how that affects chart performance.